Medical Concept Extraction

### 2012-12-16 paper/paper.tex
svms

Our system addresses the particular challenge of assigning concept labels to terms in medical records, but the task of assigning labels from multiple classes over a set of entries is a classic machine learning problem. Multi-class classification algorithms assign labels to input data entries using a model trained on a data set of pre-labeled observations. There are countless methods that support multi-class classification--for the purpose of this work, we employ three well-known strategies (Support Vector Machines, Linear Regression, and Conditional Random Fields) and compare their respective accuracies given our data sets and chosen features (section 4).

\subsubsection{Support Vector Machines}

\subsubsection{Linear Regression}
Clarification: liblinear is not doing linear regression, the default is L2-regularized SVM solver using L2-loss SVC in the dual.

@@

\subsubsection{Word Shape}
In the same spirit of our word length and MIT RegEx features, we decided to look at the high-level shapes of individual words for better generalization. More specifically, we encode words into new strings that give a sense of their capitalization and punctuation patterns as well as approximate length. We tested a variety of open-sourced word shape classification methods including two methods written by Dan Klein, one of which simply classifies words as numeric, strictly upper or lower cased, or mixed, and a second, more fine-grained classifier that attempts to group words into equivalence classes by collapsing sequences of the same time. Additionally classifiers developed by Christopher Manning seek to make less distinctions mid-sequence by sorting words based on length. Each word shape classifier result added an additional dimension to our feature space. By naively including all word shape alternatives as features, we were able to improve our testbed CRF F1 score from 0.82 to 0.85. After further testing, we settled on TODO as the most promising word shape metric because TODO.

>>>>

Our system addresses the particular challenge of assigning concept labels to terms in medical records, but the task of assigning labels from multiple classes over a set of entries is a classic machine learning problem. Supervised multi-class classification algorithms assign labels to input data entries using a model trained on a data set of pre-labeled observations. There are countless methods that support multi-class classification--for the purpose of this work, we employ three well-known strategies (general Support Vector Machines, linear Support Vector Machines, and Conditional Random Fields) and compare their respective accuracies given our data sets and chosen features (section 4).

\subsubsection{Support Vector Machines}
Support vector machines (SVMs) aim to maximize the margin between decision hyperplanes, which separate training examples in feature space. Training examples are represented as points in feature space that are mapped to create independent classification regions with the largest possible gap between regions. In use, the SVM maps new examples into this space and assigns a label based on the classification region in which the example point resides.

Often problem sets are not linearly separable in feature space. To overcome this challenge, SVMs often map the original feature space into a higher-dimensional space--the specific mapping defined by a kernel function, which is chosen for efficiency based on the specific classification problem. Separating hyperplanes, which are defined as vectors, can then be chosen as linear combinations of feature vectors in the training set.

\subsubsection{Linear SVMs}
There are some large data for which with/without nonlinear mappings gives similar performances. Without using kernels, one can quickly train a much larger set via a linear classifier.

Clarification: liblinear is not doing linear regression, the default is L2-regularized SVM solver using L2-loss SVC in the dual.

@@

\subsubsection{Word Shape}
In the same spirit of our word length and MIT RegEx features, we decided to look at the high-level shapes of individual words for better generalization. More specifically, we encode words into new strings that give a sense of their capitalization and punctuation patterns as well as approximate length. We tested a variety of open-sourced word shape classification methods including two methods written by Dan Klein, one of which simply classifies words as numeric, strictly upper or lower cased, or mixed, and a second, more fine-grained classifier that attempts to group words into equivalence classes by collapsing sequences of the same time. Additionally classifiers developed by Christopher Manning seek to make less distinctions mid-sequence by sorting words based on length. Each word shape classifier result added an additional dimension to our feature space. By naively including all word shape alternatives as features, we were able to improve our testbed CRF F1 score from 0.82 to 0.85.
